---
title: "TeenGrowth Vignette"
author: "Katherine Schaumberg "
date: "`r Sys.Date()`"
output: rmarkdown::github_document

vignette: >
  %\VignetteIndexEntry{TeenGrowth Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, warning=FALSE, message=FALSE}

knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,  # Set global figure width
  fig.height = 6  # Set global figure height
)
```

# Introduction

This vignette provides a detailed walkthrough on how to use the TeenGrowth package to forecast and visualize BMI data. This guide will cover data cleaning, BMI forecasting, and plotting the results.

```{r, echo = FALSE, message=FALSE, warning=FALSE}
library(TeenGrowth)
library(tidyr)
library(dplyr)
```

## Clean Data

The first step is to clean the data using the clean_data function. This function prepares the data by standardizing column names and units.

```{r, warning=FALSE, message=FALSE}
clean_data = clean_data(demo,
                        id_col_name = 'participant', 
                        age_col_name = 'age',
                        sex_col_name = 'sex',
                        ht_col_name = 'height',
                        wt_col_name = 'weight',
                        adult_ht_col_name = 'adult_height_in', 
                        ed_aoo_col_name = 'ed_aoo', 
                        ht_unit = 'in', 
                        wt_unit = 'lb')
```

## Create BMI Forecasts

Next, we generate BMI forecasts using the forecast_bmi function. This function calculates future BMI values based on the provided data.

```{r, warning=FALSE, message=FALSE}
BMI_forecast <- forecast_bmi(
             data = clean_data, 
             central_value = "mean",
             ci = 95)


```

## Plot Data

We can now visualize the cleaned data and the BMI forecasts. The `plot_eBMI` function is used to plot the estimated BMI data, while the `plot_weight` function plots the weight data.

```{r, warning=FALSE, message=FALSE}
plot_eBMI(clean_data = clean_data, 
          forecast_data = BMI_forecast, 
          px = 2)
```

```{r, warning=FALSE, message=FALSE}
plot_weight(clean_data = clean_data, 
          forecast_data = BMI_forecast, 
          px = 2)

```

# 10-year BMI and Weight Forecasts

An additional function, `clean_forecast_data` allows for the creation of a table that organizes 10-year forecast data (in 6-month age intervals) for expected BMI and Weight

```{r, warning=FALSE, message=FALSE}
forecast_data <- clean_forecast_data(BMI_forecast, 
                    px = 2, 
                    model = 'mean')
knitr::kable(forecast_data)

```

# Weight Restoration Planning

Finally, we provide an example of how to use the package for treatment planning. In this case, a specific participant is identified, then the data is prepared for plotting with `tx_plot_clean`, and then a weight restoration plot is able to be derived using the `Wt_Restore_Plot` function.

```{r, warning=FALSE, message=FALSE}

wt_restore <- demo |> filter(participant == 2)
wt_restore_clean <-     tx_plot_clean(wt_restore,
                        age_col_name = 'age',
                        age_unit = 'years',
                        ht_col_name = 'height',
                        wt_col_name = 'weight',
                        adult_ht = wt_restore$adult_height_in[1],
                        ht_unit = 'in', 
                        wt_unit = 'lb', 
                        dob = '2009-01-01', 
                        tx_start_date = '2024-01-01')

  
wt_restore_forecast <- BMI_forecast |> filter (id == 2)
  

Wt_Restore_Plot(wt_restore_clean, 
                wt_restore_forecast, 
                slope_per_week = 0.5)

```
